An Embedding Framework for Consistent Polyhedral Surrogates

### 2021-06-07 poly-embed-journal/poly-embed-journal.tex
Re-do Prop 7 proof with new Lemma on weakly dominated reports

\begin{restatable}{proposition}{embedtrim}\label{prop:embed-trim}
  Let $\Gamma:\simplex\toto\reals^d$ be an elicitable property.
  The following are equivalent:
  \begin{enumerate}\setlength{\itemsep}{0pt}
  \item $\Gamma$ embeds a elicitable finite property $\gamma:\simplex \toto \R$.
  \item $\trimnonred(\Gamma)$ is a finite set, and $\cup\,\trimnonred(\Gamma) = \simplex$.
  % \\ $\hat\Gamma := \trimnonred(\Gamma)$ is a finite property, and $\cup\,\hat\Gamma) = \simplex$.
  \item There is a finite minimum representative set $\U$ for $\Gamma$.
  \item There is a finite set of full-dimensional level sets $\Theta$ of $\Gamma$, and $\cup\,\Theta = \simplex$.
  \end{enumerate}
  % \raft{I see why you wanted this addition here, but (a) it's clear from the def of embedding, and (b) we don't have varphi defined.  So let's leave it out.}
  Moreover, when any of the above hold, $\trimnonred(\gamma) = \trimnonred(\Gamma) = \{\Gamma_u : u\in\U\} = \Theta$.\jessiet{Also, should this statement be $\trimcover(\gamma) = \trimnonred(\gamma) = \trimcover(\Gamma) = \trimnonred(\Gamma) = \{\Gamma_u \mid u \in \U\} = \Theta$? But also, can we pull it out into its own statement? It should follow as a corollary of this and an analogous version of Prop 6.}
\end{restatable}
\begin{proof}
	\jessie{Keep plus example in mind}
	\jessie{Use trim nonredundant and trim covering macros}
	\jessie{1 -> 2 -> 3 -> 1, 3 <=> 4}
	Let $L$ elicit $\Gamma$.

	1 $\Rightarrow$ 2:
	%
	% \jessie{Loss intuition: Let $L$ elicit $\Gamma$.  If $\Gamma$ embeds a finite $\gamma'$, $L$ must embed a discrete loss $\ell$ such that $\ell(r) = L(\varphi(r))$.  Therefore, $\risk{L} = \risk{\ell}$ is polyhedral. Moreover, we have $\Gamma$ also embedding the finite property $\prop{\ell} := \gamma$ as a corollary of the definition of embedding properties. We claim $\trim(\Gamma) = \{\gamma_r : r \in \R\}$ (Lemma~\ref{lem:embedded-level-sets-trim}), and since the latter set is finitely generated (as $\R$ is finite), we have have $\trim(\Gamma)$ is finite.  Moreover, as $\gamma$ is nondegenerate (since it is defined over the simplex), we have that $\cup_{r \in \R} \gamma_r = \simplex$.}
	%
%	Since $\gamma$ is finite, it has a finite minimum representative set $\Sc$, and we can show $\trimnonred(\gamma) = \trimcover(\gamma) := \{\gamma_r \mid r \in \Sc\}$. 
%	\jessiet{Hand waving}
%	Moreover, since the embedding condition implies $\gamma_r = \Gamma_{\varphi(r)}$, we also have $\trimcover(\Gamma) = \trimcover(\gamma)$, which is finite as $\gamma$ is finite.
%	Now we can show that $\trimcover(\Gamma)$ finite implies $\trimcover(\Gamma) = \trimnonred(\Gamma)$ and the result follows.
%
	%Since $\Gamma$ embeds a finite $\gamma$ over some minimum representative set $\Sc$, and each for each $r \in \Sc$, we have $\gamma_r = \Gamma_{\varphi(r)}$.
	%
	Since $\R$ is finite, we have $\trimnonred(\Gamma) \subseteq \{\Gamma_{\varphi(r)} \mid r \in \R\} = \{\gamma_r \mid r \in \R\}$.
	As the last set is finite, so is $\trimnonred(\Gamma)$.
	Moreover, as $\trimnonred$ only removes redundant reports, we have $\cup \trimnonred(\Gamma) = \simplex$.

\btw{Commenting out 3 June 21 to use old $\trimnonred$ definition.}
%	If $\Gamma$ embeds a finite $\gamma$, then any representative set $\Sc \subseteq \R$ is finite, and so any minimum representative set is finite; $\trim$ is then finite as it is finitely generated.
%	$\cup \trim(\Gamma) = \simplex$ follows from representativeness.

\btw{Commenting out 2 $\Rightarrow$ 1 for now since I added 1-> 2 -> 3 ->1 for cyclicity.}
%	\raft{These next two directions together seem like what we did before for $2\Rightarrow 3$, correct?  (Not as complete though; see the step where we justify the conditions of Lemma~\ref{lem:loss-restrict}.)  I'm actually happy for them to be separated, but let's make sure the argument is just as complete. \jessie{Added in the conditions of Lemma 3 to be a bit more thorough.}}\btw{Not sure how old this comment is, but feel free to burn if you're happy with thoroughness.}
%	2 $\Rightarrow$ 1:
%	As $\trim(\Gamma)$ is finite, we know $\trim(\Gamma) = \{\Gamma_{u_i} \mid u_i \in \U\}$ for some minimum representative set $\U$ of finite cardinality $k$.
%	Then we can apply Lemma~\ref{lem:loss-restrict} to observe that $L$ embeds $L|_\U$, and therefore $\Gamma$ embeds $\prop{\ell}$, which is finite as $\ell$ is discrete.

	%	1 $\Rightarrow$ 3: \jessie{Intuition: Each $\theta \in \trim(\Gamma)$ is full dimensional by Lemma~\ref{lem:embedded-level-sets-trim} and equality of the set of embedded level sets of $\Gamma$ and set of level sets of $\gamma$.  Take $\Theta = \trim(\Gamma)$.}
	%	Assume $\Gamma$ embeds a finite property $\gamma$.
	%	Then for each $r \in \R$, we have $\Gamma_{\varphi(r)} = \gamma_r$, and therefore $\{\Gamma_{\varphi(r)} : r \in \R\} = \{\gamma_r : r \in \R\}$.
	%	Chaining this with Lemma~\ref{lem:embedded-level-sets-trim}, we have $\trim(\Gamma) = \{\gamma_r : r \in \R\}$.
	%	
	%	\jessie{This is really the beginning of 2 $\Rightarrow 3$}
	%	If there is a level set $\Gamma_{\varphi(\hat r)} \in \trim(\Gamma)$ that is not full-dimensional, then we can define an injection $\varphi'$ so that $\varphi' :\R \setminus \{\hat r\} \to \reals^d$ maps $r \mapsto \varphi(r)$.
	%	(Iterate if there are multiple level sets that are not full-dimensional.)
	%	Now take $\Theta = \{\Gamma_{\varphi'(r)} : r \in \R\}$.
	%	We know each level set in $\Theta$ is full-dimensional by construction, and just need to show $\cup \Theta = \simplex$.
	%	However, if there was a level set $\Gamma_{\varphi(\hat r)}$ that was not in $\trim(\Gamma)$ \jessie{\ldots}

	2 $\Rightarrow$ 3: 
	%Suppose $\trimcover(\Gamma) = \{\Gamma_u \mid u \in \U\}$.
	%Then we know $\U$ is a minimum representative set, and since $\trimcover(\Gamma)$ is finite, so is $\U$; if it was not, then $\U$ would not be minimum.
	\jessie{Using the old definition of $\trim := \trimnonred$, mirroring ish previous proofs} 
	As a reminder, we have $\trimnonred(\Gamma) = \{\Gamma_u \mid \neg \exists u' \neq u $ s.t. $ \Gamma_u \subsetneq \Gamma_{u'}\}$.
	Take $\U^* = \{u_1, \ldots, u_k\}$ to be a finite set of reports forming $\trimnonred$.
	Since $\cup \trimnonred(\Gamma) = \cup_{u \in \U^*} \Gamma_u = \simplex$, we have $\U^*$ representative.
	We can now construct a minimum representative set $\U \subseteq \U^*$.
	Form $\U$ by iterating over $\U^*$ and removing any element $u'$ of $\U^*$ if there is another $u$ remaining in $\U^*$ such that $\Gamma_u = \Gamma_{u'}$.
	As we only remove exact duplicate level sets, we still have $\U$ representative, and proceed to show $\U$ is minimum.

	If $\U$ was not minimum, there would be some report $u' \in \U$ so that $\U' := \U \setminus \{u'\}$ is also representative.
	Since $\Gamma$ is elicitable and $\trimnonred$ is finite, we know that the level sets of $\Gamma$ form a power diagram intersected with the simplex (Theorem~\ref{thm:aurenhammer}).
	Therefore, every level set $\Gamma_u$ corresponds to a face of a convex polyhedron: we correspond in two cases, according to if $\Gamma_{u'}$ corresponds to a facet or some lower-dimensional face.

	First, suppose $\Gamma_{u'}$ does not correspond to a facet.
	Since $\Gamma_{u'}$ does correspond to a face of a convex polytope, it must be contained in a facet, and therefore there is some $u$ such that $\Gamma_{u'} \subsetneq \Gamma_u$.
	Thus, $\Gamma_u \not \in \trimnonred(\Gamma)$.
	If $\Gamma_{u'}$ corresponds to a facet, then it has a well-defined nonempty interior, and by removing exactly duplicate reports in the construction of $\U$, we have a $p \in \inter{\Gamma_{u'}}$ such that $\{u'\} = \Gamma(p)$.
	Therefore, $\U'$ would not be representative.
	We conclude $\U$ as constructed must be a minimum representative set.

	3 $\Rightarrow$ 1:
	Since $\U$ is representative, we have $\Gamma|_\U(p) := \Gamma(p) \cap \U$ nonempty for all $p \in \simplex$, and therefore, $\Gamma$ embeds $\Gamma|_\U$, which is finite as $\U$ is finite.
	Since $\U$ is minimum, the constructed $\Gamma_\U$ is additionally non-redundant.
	\jessiet{Want more details here.  It's the forward implication of Prop 4, but since this is used in the reverse implication of prop 4, don't want to create cyclic dependencies.}

	\btw{Previous version commented out above because it relies on old definitions - Jessie 1 June 21, commenting out 1 $\iff$ 4 to replace with 3 $\iff$ 4 - 3 June 21}
%	1 $\Rightarrow$ 4:
%	Let $\varphi: \R \to \reals^d$ be the embedding function, and take $\Theta = \{\Gamma_{\varphi(r)} : r \in \R\}$.
%	The latter set is equal to $\{\gamma_r : r \in \R\}$ by Lemma~\ref{lem:embedded-level-sets-trim}.
%	Therefore, we know each $\theta \in \Theta = \gamma_r$ for some $r \in \R$, which is full-dimensional in the affine hull of the simplex.
%	Moreover, since $\gamma$ is non-degenerate, we have $\cup \Theta = \cup_r \gamma_r = \simplex$.	
%	Suppose $\trimcover(\Gamma) = \{\Gamma_u \mid u \in \U\} = \simplex$.
%	(Embedding implies $\Sc$ is a minimum representative set for $\gamma$ iff $\varphi(\Sc)$ is minimum representative for $\Gamma$.)
%	Take $\Theta = \trimcover(\Gamma)$; since $\U$ is representative, we know $\cup \Theta = \simplex$.
%	Moreover, $\Theta$ is finite since $\U$ is finite.
%	It is just left to show each $\theta \in \Theta$ is full-dimensional.
%	Since $\Gamma$ is elicitable and has a finite minimum representative set, 
%	Since $\gamma$ is a finite property, its level sets form a power diagram in the simplex (Theorem~\ref{thm:aurenhammer}).
%	If there was a $\theta = \gamma_r$ that was not full-dimensional, it must therefore be the face of another level set.
	%Since level sets of elicitable properties are closed
	%\footnote{$\gamma_r = \Gamma_{\varphi(r)}$ for all $r \in \Sc$, we can use this characteristic about level sets of elicitable properties. \jessie{Adding this since we haven't shown $\gamma$ elicitable yet.}}, 
	%$\gamma_r$ must then be a face of some facet $\gamma_{r'}$.
	%Therefore, $\gamma_r \subseteq \gamma_{r'}$, contradicting $\Sc$ being minimum, so each $\theta \in \Theta$ must be full-dimensional.
%	%	Since $\Gamma$ embeds the finite property $\gamma$, Lemma~\ref{lem:embed-defs-equiv} yields a loss $\ell$ eliciting $\gamma$ embedded by $L$, and therefore $\risk{L} = \risk{\ell}$.
%	%	As in the proof of Proposition~\ref{prop:embed-bayes-risks}, we observe that every facet of the epigraph of $-\risk{\ell}$ projects onto the simplex as a full-dimensional level set of $\gamma$.
%	%	As the risks are equal, this is also true for $\Gamma$.
%	%	Observe each cell in the simplex of $\gamma$ has some report $r$ so the cell is $\gamma_r$.
%	%	If $\varphi: r \mapsto u_r$, then we must have $\gamma_r = \Gamma_{u_r}$.
%	%	Take $\Theta$ to be the set $\{\Gamma_{u_r} : u_r \in \varphi(\R)\}$.
%	%	We have shown each $\Gamma_{u_r}$ is full-dimensional, and as the union of these cells forms a power diagram on $\simplex$, we have $\cup \Theta = \simplex$.
%	%	2 $\Rightarrow$ 3:
%	%	\jessie{New attempt: show that $\Gamma$ elicitable and $\trim(\Gamma)$ finite implies each $\Gamma_u \in \trim(\Gamma)$ is full-dimensional. }
%	%	We claim every $\theta \in \trim(\Gamma)$ is full-dimensional, then take $\Theta = \trim(\Gamma)$ and the result holds.
%	%	Suppose there was a $u \in \reals^d$ so that $\Gamma_u \in \trim(\Gamma)$ and $\Gamma_u$ was not full-dimensional.
%	%	The only case in which this happens is if $\Gamma_u$ spans two level sets, as anything that did not would be a subset of a full-dimensional level set in $\trim(\Gamma)$.
%	%	However, if $\Gamma_u$ spans the relative interior of two full-dimensional level sets, we contradict the elicitability condition in Lemma~\ref{lem:elicitable-level-sets}.\jessiet{Need to be more careful here.}
%	%	(If $\Gamma_u$ spans only the boundary of one level set, then we have that $\Gamma_u$ is a subset of the full-dimensional level set by closure of level sets of elicitable properties.)
%	%	Therefore, we must have each $\Gamma_u \in \trim(\Gamma)$ full-dimensional if $\trim(\Gamma)$ is a finite set and $\Gamma$ is elicitable.
%	4 $\Rightarrow$ 1: 
%	Let $\Theta = \{\theta_1,\ldots,\theta_k\}$.
%	For all $i\in\{1,\ldots,k\}$ let $u_i\in\reals^d$ such that $\Gamma_{u_i} = \theta_i$.
%	Now define $\gamma:\simplex\toto\{1,\ldots,k\}$ by $\gamma(p) = \{i : p\in\theta_i\}$, which is non-degenerate as $\cup\,\Theta = \simplex$.
%	By construction, we have $\gamma_i = \theta_i = \Gamma_{u_i}$ for all $i$, so letting $\varphi(i) = u_i$ we satisfy the definition of embedding, namely statement (i.).

3 $\Rightarrow$ 4:
It suffices to show that, for each $u \in \U$, $\Gamma_u$ is full-dimensional, so we can take $\Theta = \{\Gamma_u \mid u \in \U\}$.
Since $\Gamma$ has a finite representative set, we know its level sets are faces of convex polyhedron by equivalence to a power diagram intersected with the simplex.
If there is a $u \in \U$ such that $\Gamma_u$ is not full-dimensional, it must be a face of some facet corresponding to $\Gamma_{u'}$ for some $u' \in \U$, and hence we contradict $\U$ being minimum, as $\U \setminus \{u\}$ would be a representative set of strictly smaller cardinality.

4 $\Rightarrow$ 3:
We want to show that for each $\theta \in \Theta$, we cannot have $\Theta' := \Theta \setminus \{\theta\}$ be representative.
Observe we have no duplicate elements of $\Theta$.
Fix $\theta := \Gamma_u \in \Theta$.
Since $\Gamma_u$ is full-dimensional, there is some $p \in \inter{\Gamma_u}$ so that $\{u\} = \Gamma(p)$.
Therefore, $\Theta'$ is not representative, so $\Theta$ is a finite minimum representative set for $\Gamma$.

Observe now that when one of these is true, $\Gamma|_\U$ is elicited by $L|_\U$, and that there is a bijection from $\Gamma|_\U$ to $\gamma$, and therefore $\gamma$ is also elicitable. 
\end{proof}

>>>>

\begin{lemma}\label{lem:finite-minrep-weakly-dominate}
	Let $\Gamma : \simplex \toto \R$ be an elicitable property with a finite minimum representative set $\R^*$.
	For any $r \in \R \setminus \R^*$, there is a $r^* \in \R^*$ such that $\Gamma_r \subseteq \Gamma_{r^*}$.
\end{lemma}
\begin{proof}
	Let $L$ elicit $\Gamma$.
	By Proposition~\ref{prop:SL-minimum}, since $L$ has a finite minimum representative set, $-\risk{L}$ is polyhedral, with epigraph $E_L$.
	Observe that for any report $r \in \R$, the hyperplane $H^L_r : p \mapsto (p, -\inprod{p}{L(r)})$ supports $E_L$ on exactly the level set.
	Since $E_L$ is a convex polyhedron in $\reals^{n+1}$, we can take a linear projection of the faces of $E_L$ onto $\simplex \subset \reals^n$ to yield an affinely equivalent cell complex $C$ (Definition~\ref{def:cell-complex}), which is a power diagram by Theorem~\ref{thm:aurenhammer}.
	Thus, every level set of $\Gamma$ is a cell in the induced power diagram.

	Observe that any $r^* \in \R^*$ must be a full-dimensional cell of the power diagram, since if it was a lower-dimensional, it would be a subset of some full-dimensional cell, and we would contradict that $\R^*$ is minimum.
	Moreover, if any $r \in \R \setminus \R^*$ is full-dimensional, $\Gamma_r$ is exactly equal to $\Gamma_{r^*}$ for some $r^* \in \R^*$ since $\R^*$ is representative.
	If $\Gamma_r$ is not full-dimensional, it must be a face of some full-dimensional level set $\Gamma_{r^*}$, and therefore $\Gamma_r \subset \Gamma_{r^*}$.
\end{proof}

\begin{restatable}{proposition}{embedtrim}\label{prop:embed-trim}
  Let $\Gamma:\simplex\toto\reals^d$ be an elicitable property.
  The following are equivalent:
  \begin{enumerate}\setlength{\itemsep}{0pt}
  \item $\Gamma$ embeds a elicitable finite property $\gamma:\simplex \toto \R$.
  \item $\trimnonred(\Gamma)$ is a finite set, and $\cup\,\trimnonred(\Gamma) = \simplex$.
  % \\ $\hat\Gamma := \trimnonred(\Gamma)$ is a finite property, and $\cup\,\hat\Gamma) = \simplex$.
  \item There is a finite minimum representative set $\U$ for $\Gamma$.
  \item There is a finite set of full-dimensional level sets $\Theta$ of $\Gamma$, and $\cup\,\Theta = \simplex$.
  \end{enumerate}
  % \raft{I see why you wanted this addition here, but (a) it's clear from the def of embedding, and (b) we don't have varphi defined.  So let's leave it out.}
  Moreover, when any of the above hold, $\trimnonred(\gamma) = \trimnonred(\Gamma) = \{\Gamma_u \mid u\in\U\} = \Theta$.
\end{restatable}
\btw{Deleted old proof (too messy for me to track) 7 June 21 at 8:47 AM}
\begin{proof}
Let $L$ elicit $\Gamma$, and $\ell$ elicit $\gamma$ if given.

$1 \Rightarrow 2, 3$:
Since $\Gamma$ embeds a finite property $\gamma$ with minimum representative set $\Sc$, $\Gamma$ also has a finite minimum representative set $\U := \varphi(\Sc)$. (3).

For (2), we claim $\trimnonred(\Gamma) = \{\Gamma_u \mid u \in \U\}$, and the result follows from $\U$ being a finite minimum representative set.
Since $\U$ is a finite minimum representative set, $\trimnonred(\Gamma) \supseteq \{\Gamma_u \mid u \in \U\}$ follows immediately.
We now show $\trimnonred(\Gamma) \subseteq \{\Gamma_u \mid u \in \U\}$.
Consider any $\Gamma_{u'} \in \trimnonred(\Gamma)$.
Since $\Gamma$ has the finite minimum representative set $\U$, if $u' \not \in \U$, there is some $u \in \U$ such that $\Gamma_{u'} \subseteq \Gamma_u$ by Lemma~\ref{lem:finite-minrep-weakly-dominate}. 
If $\Gamma_{u'} \subsetneq \Gamma_u$, then $u' \not \in \U_\equiv$, and $\Gamma_{u'} \not \in \trimnonred(\Gamma)$ nor is $u' \in \U$.
If $\Gamma_{u'} = \Gamma_u$, then$u' \equiv u$ so $\Gamma_{u'} \in \trimnonred(\Gamma)$ as $[u'] = [u] \in \U_\equiv$.
Thus, we have $\trimnonred(\Gamma) = \{\Gamma_u \mid u \in \U\}$, so $\trimnonred(\Gamma)$ is finite and unions to the simplex.

%$2 \Rightarrow 3$:
$3 \Rightarrow 1$:
Consider the loss $\ell := L|_\U$.
$\ell$ is a finite loss since $\U$ is finite, and by Corollary~\ref{cor:representative-embeds-restriction}, $L$ embeds $\ell$.
As a corollary of the loss embedding condition, we have $\Gamma$ embeds $\gamma := \prop{\ell}$.

$3 \Rightarrow 4$:
Since level sets of elicitable properties are closed, the set $\Theta = \{\Gamma_u \mid u \in \U\}$ yields a finite covering of the simplex.\jessiet{How much effort do we need to put into showing each level set is full-dimensional?}

$4 \Leftarrow 3$:
Since $\Theta$ is a finite set of full-dimensional level sets, and these level sets are affinely equivalent to the epigraph of negative risk $\mathrm{epi}(-\risk L)$, it must follow that $-\risk L$ is polyhedral (convexity follows as a property of $-\risk L$). 
By Proposition~\ref{prop:SL-minimum}, $L$ has polyhedral risk if and only if $L$ (and therefore $\Gamma$) has a finite minimum representative set $\U$.
From Proposition~\ref{prop:SL-minimum}, we also know there is a bijection between the facets of $\mathrm{epi}(-\risk L)$ (corresponding to the level sets $\theta \in \Theta$) and the level sets $\{\Gamma_u \mid u \in \U\}$, so the generated sets are equal.
